Multi-Robot Probabilistic Mapping and Exploration

### 2018-06-26 Dissertation/Thesis/tex/ae2D.tex
edited up to sec. 4.2

Dijkstra's search is naturally applied to this problem because the occupancy grid map serves nicely as a graph, and Dijkstra's algorithm provides an optimal collision-free trajectory along that graph. There are two steps to Dijkstra's algorithm. first, generate a cost map from the robot location to each grid cell on the 2D occupancy grid map. Only cells that satisfy \refeqn{CollisionInequalityConstraint} are considered to avoid possible collisions. The cost to travel to neighboring cells that share an edge are the cell edge length $\alpha$, and the cost to travel to cells that share a corner is $\sqrt{2}\alpha$. This provides the collision-free distances to all reachable locations on the map. Second, the waypoints from a candidate pose to the current robot pose are easily obtained along the cost map using steepest descent.

@@

Here, a robot autonomously maps and explores a 2D environment composed of two rooms and one hallway. The robot models its surroundings with an occupancy grid of $15000$ cells where grid cell edges have length $\alpha=0.2$m, composing a map with dimensions $30\text{m}\times20\text{m}$. The initial probability $P(\mathbf{m}_i)=1\times10^{-10}\approx0$ (minimum value for free space) for grid cells covered by the circular robot of radius $0.1$m and $P(\mathbf{m}_i)=0.5$ for all other cells.
At each time step, the robot receives a measurement scan, where the probabilistic properties of the sensor are taken from~\cite{PirRutBisSch11,KhoElb12}. Then, the $n_c=8$ evenly-spaced candidate locations about a circle of radius $\delta=0.5$m around the current pose location are considered, where $n_d=32$ measurement rays are evenly-spaced about the candidate location.  When no current candidates yield expected information gains above $\mathcal I_\text{min}=2$, $\lambda=1.25$ is multiplied to $n_c$ and $\delta$. The motion of the robot is restricted to movement along cells satisfying \refeqn{CollisionInequalityConstraint} with $\beta=0.01$. Dijkstra's algorithm provides collision-free motion planning.
The results are illustrated in Figure \ref{fig:AutonomousExploration}.

@@

Knowing only that the robot is inside free space at the beginning, the robot carefully navigates the environment while avoiding collisions. The robot motion is governed by a policy that maximizes the map information gain within its set of pose candidates, where the $\hat n=6$ is chosen to approximate the expected entropy of each ray. 
When running the exploration algorithm in Robot Operating System (ROS), the mean computation time is $0.0194$ seconds to determine the optimal future pose and complete Dijkstra's algorithm on the occupancy grid.
The computation times for this map reached roughly $1$ second at maximum, corresponding to rare cases when the robot is locally surrounded by a highly-certain environment; this task requires evaluating many more pose candidates and motion planning over a larger terrain.

@@

Next, the proposed autonomous exploration approach is applied to the floor plan of Intel Research Lab, illustrated in Figure~\ref{fig:intel}.
The robot explores its surroundings with an occupancy grid with $90,000$ cells where grid cell edges are $\alpha=0.1$m, composing a map with dimensions $30\text{m}\times30\text{m}$.
Similar to the prior example, the initial probability $P(\mathbf{m}_i)=1\times10^{-10}\approx0$ (minimum value for free space) is chosen for grid cells covered by the circular robot of radius of $0.3$m and $P(\mathbf{m}_i)=0.5$ for all other cells.
For added safety, cells falling within $0.6$m of the robot are considered inside the possible collision zone $\mathcal{C}_X$, from \refeqn{CollisionInequalityConstraint}, where $\beta=0.1$ is selected.

@@

The time sequence of the exploration for every 5~minutes are simulated in ROS Stage 2D environment. The corresponding simulation results are illustrated in Figure \ref{fig:IRL} and a video is available at \href{https://www.youtube.com/watch?v=5VdzKHreB_s}{\WriteBlue{https://www.youtube.com/watch?v=5VdzKHreB\_s}}.
Lighter red dots and darker blue dots in the explored area correlate to the level of information gain at those positions with low and high value, respectively. It is shown that the proposed autonomous exploration algorithm successfully mapped the complicated environment composed of a number of rooms, narrow hallways, and open spaces that are irregularly shaped.

\begin{figure}
    \centering
    \includegraphics[width=0.4\textwidth]{intel_clean.png}
    \caption{Modified Intel Research Lab floor plan form the SLAM benchmark~\cite{kummerle2009measuring}.}
\label{fig:intel}
\end{figure}

\begin{figure}[!ht]
    \centering
    \begin{subfigure}[t]{0.3\columnwidth}
        \centering
        \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{0min.png}
        \caption{$t=0$ min}
        \label{fig:IRL0min}
    \end{subfigure}
    \begin{subfigure}[t]{0.3\columnwidth}
        \centering
        \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{5min.png}
        \caption{$t=5$ min}
        \label{fig:IRL5min}
    \end{subfigure}
    \begin{subfigure}[t]{0.3\columnwidth}
           \centering
           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{10min.png}
        \caption{$t=10$ min}
        \label{fig:IRL10min}
    \end{subfigure}
    \begin{subfigure}[t]{0.3\columnwidth}
           \centering
           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{15min.png}
        \caption{$t=15$ min}
        \label{fig:IRL15min}
    \end{subfigure}
    \begin{subfigure}[t]{0.3\columnwidth}
         \centering
         \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{20min.png}
        \caption{$t=20$ min}
        \label{fig:IRL20min}
    \end{subfigure}
    \begin{subfigure}[t]{0.3\columnwidth}
           \centering
           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{25min.png}
        \caption{$t=25$ min}
        \label{fig:IRL25min}
    \end{subfigure}
    \begin{subfigure}[t]{0.3\columnwidth}
           \centering
           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{30min.png}
        \caption{$t=30$ min}
        \label{fig:IRL30min}
    \end{subfigure}
    \begin{subfigure}[t]{0.3\columnwidth}
           \centering
           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{35min.png}
        \caption{$t=35$ min}
        \label{fig:IRL35min}
    \end{subfigure}
    \begin{subfigure}[t]{0.3\columnwidth}
           \centering
           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{40min.png}
        \caption{$t=40$ min}
        \label{fig:IRL40min}
    \end{subfigure}
    \caption{A robot (green) measures a room in ROS Stage simulator using the Intel Research Lab floor plan. The exact inverse sensor model is used for the autonomous exploration algorithm.}
    \label{fig:IRL}
\end{figure}

>>>>

Dijkstra's search is naturally applied to this problem because the occupancy grid map serves nicely as a graph, and Dijkstra's algorithm provides an optimal collision-free trajectory along that graph. There are two steps to Dijkstra's algorithm: first, generate a cost map from the robot location to each grid cell on the 2D occupancy grid map. Only cells that satisfy \refeqn{CollisionInequalityConstraint} are considered to avoid possible collisions. Traveling to a neighboring cells that shares an edge costs the edge length distance $\alpha$, and the cost to travel to cells that share a corner is $\sqrt{2}\alpha$. This provides the collision-free distances to all reachable locations on the map. Second, the waypoints from a candidate pose to the current robot pose are easily obtained along the cost map using steepest descent.

@@

Here, a robot autonomously maps and explores a 2D environment composed of two rooms and one hallway. The robot models its surroundings with an occupancy grid of $15,000$ cells where grid cell edges have length $\alpha=0.2$m, composing a map with dimensions $30\text{m}\times20\text{m}$. The initial probability $P(\mathbf{m}_i)=1\times10^{-10}\approx0$ (minimum value for free space) for grid cells covered by the circular robot of radius $0.1$m and $P(\mathbf{m}_i)=0.5$ for all other cells.
At each time step, the robot receives a measurement scan, where the probabilistic properties of the sensor are taken from~\cite{PirRutBisSch11,KhoElb12}. Then, the $n_c=8$ evenly-spaced candidate locations about a circle of radius $\delta=0.5$m around the current pose location are considered, where $n_d=32$ measurement rays are evenly-spaced about the candidate location.  When no current candidates yield expected information gains above $\mathcal I_\text{min}=2$, $\lambda=1.25$ is multiplied to $n_c$ and $\delta$. The motion of the robot is restricted to movement along cells satisfying \refeqn{CollisionInequalityConstraint} with $\beta=0.01$. Dijkstra's algorithm provides collision-free motion planning.
The results are illustrated in Figure \ref{fig:AutonomousExploration}.

@@

Knowing only that the robot is inside free space at the beginning, the robot carefully navigates the environment while avoiding collisions. The robot motion is governed by a policy that maximizes the map information gain within its set of pose candidates, where the $\hat n=6$ is chosen to approximate the expected entropy of each ray. 
When running the exploration algorithm within the framework of the Robot Operating System (ROS), the mean computation time is $0.0194$ seconds to determine the optimal future pose and complete Dijkstra's algorithm on the occupancy grid.
The computation times for this map reached roughly $1$ second at maximum, corresponding to rare cases when the robot is locally surrounded by a highly-certain environment; this task requires evaluating many more pose candidates and motion planning over a larger terrain.

@@

Next, the proposed autonomous exploration approach is applied to the floor plan of the Intel Research Lab, illustrated in Figure~\ref{fig:intel}.
The robot explores its surroundings with an occupancy grid with $90,000$ cells where grid cell edges are $\alpha=0.1$m, composing a map with dimensions $30\text{m}\times30\text{m}$.
Similar to the prior example, the initial probability $P(\mathbf{m}_i)=1\times10^{-10}\approx0$ (minimum value for free space) is chosen for grid cells covered by the circular robot of radius of $0.3$m and $P(\mathbf{m}_i)=0.5$ for all other cells.
For added safety, cells falling within $0.6$m of the robot are considered inside the possible collision zone $\mathcal{C}_X$, from \refeqn{CollisionInequalityConstraint}, where $\beta=0.1$ is selected.

@@

The time sequence of the exploration for every 5~minutes are simulated in the ROS Stage 2D environment. The corresponding simulation results are illustrated in Figure \ref{fig:IRL} and a video is available at \href{https://www.youtube.com/watch?v=5VdzKHreB_s}{\WriteBlue{https://www.youtube.com/watch?v=5VdzKHreB\_s}}.
Lighter red dots and darker blue dots in the explored area correlate to the level of information gain at those positions with low and high value, respectively. It is shown that the proposed autonomous exploration algorithm successfully mapped the complicated environment composed of a number of rooms, narrow hallways, and open spaces that are irregularly shaped.

\begin{figure}
    \centering
    \includegraphics[width=0.4\textwidth]{intel_clean.png}
    \caption{Intel Research Lab floor plan benchmark~\cite{kummerle2009measuring}.}
\label{fig:intel}
\end{figure}

\begin{figure}[!ht]
    \centering
    \begin{subfigure}[t]{0.3\columnwidth}
        \centering
        \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{0min.png}
        \caption{$t=0$ min}
        \label{fig:IRL0min}
    \end{subfigure}
    \begin{subfigure}[t]{0.3\columnwidth}
        \centering
        \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{5min.png}
        \caption{$t=5$ min}
        \label{fig:IRL5min}
    \end{subfigure}
    \begin{subfigure}[t]{0.3\columnwidth}
           \centering
           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{10min.png}
        \caption{$t=10$ min}
        \label{fig:IRL10min}
    \end{subfigure}
    \begin{subfigure}[t]{0.3\columnwidth}
           \centering
           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{15min.png}
        \caption{$t=15$ min}
        \label{fig:IRL15min}
    \end{subfigure}
    \begin{subfigure}[t]{0.3\columnwidth}
         \centering
         \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{20min.png}
        \caption{$t=20$ min}
        \label{fig:IRL20min}
    \end{subfigure}
    \begin{subfigure}[t]{0.3\columnwidth}
           \centering
           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{25min.png}
        \caption{$t=25$ min}
        \label{fig:IRL25min}
    \end{subfigure}
    \begin{subfigure}[t]{0.3\columnwidth}
           \centering
           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{30min.png}
        \caption{$t=30$ min}
        \label{fig:IRL30min}
    \end{subfigure}
    \begin{subfigure}[t]{0.3\columnwidth}
           \centering
           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{35min.png}
        \caption{$t=35$ min}
        \label{fig:IRL35min}
    \end{subfigure}
    \begin{subfigure}[t]{0.3\columnwidth}
           \centering
           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{40min.png}
        \caption{$t=40$ min}
        \label{fig:IRL40min}
    \end{subfigure}
    \caption{A robot (green) measures a room in the ROS Stage simulator using the Intel Research Lab floor plan. The exact inverse sensor model is used for the autonomous exploration algorithm.}
    \label{fig:IRL}
\end{figure}

### 2018-06-26 Dissertation/Thesis/tex/ae3Dsimple.tex
edited up to sec. 4.2

In this chapter, we consider mapping and exploring in 3D environments. We consider common spaces like rooms and hallways where objects are largely vertically-uniform. A complete 3D probabilistic map is projected onto a simpler 2D map at a fixed exploration height, where the projected map can be used for collision-avoidance and entropy predictions.

@@

However, implementation in 3D requires several careful considerations. First, the number of cell occupancy probabilities tracked with computer memory increases by a factor of the number of cells in the added dimension. In its current form, the proposed method assumes fixed map limits, so these should be carefully selected with grid cell resolution based on available memory. Second, measurement rays spanning 3D space typically involve more intersections with grid cells. The computational order of ray casting and the inverse sensor model are proportional to the number of grid cells along the ray $n_r$, so careful consideration must be placed on sensor limits and rays per scan that can be considered. Typically, computers small enough to be carried onboard flying robots lack the processing capabilities to update massive maps quickly, particularly because the ray-by-ray updating method requires many computations in series with large measurement scans. Therefore, an onboard computer can simply stream the sensor data, and a more powerful computer can run the mapping.

Additionally, mobile robots, aerial vehicles in particular, are frequently subject to fast movements, and the time stamps for pose estimates and depth sensor readings are not guaranteed to align. For example, the most recent pose estimate and measurement scan may have significantly different time stamps while the robot is rapidly rotating. Using these together violates the assumption that the robot pose is well-known for every measurement scan, i.e., $X_t$ and $Z_t$ are given. Therefore, pose estimates and sensor readings must be paired according to their time stamps. A useful tool that easily satisfies this goal is the Message Filters package with the Robot Operating System (ROS), where stamped poses and measurement scans are synchronized using recently received data.

>>>>

In this chapter, we introduce mapping and autonomous exploration in 3D environments. Here, we consider common spaces like rooms and hallways where objects are largely vertically-uniform. A complete 3D probabilistic map is projected onto a simpler 2D map at a fixed exploration height, where the projected map can be used for collision-avoidance and entropy predictions.

@@

However, implementation in 3D requires several careful considerations. First, the number of cell occupancy probabilities tracked with computer memory increases by a factor of the number of cells in the added dimension. In its current form, the proposed method assumes fixed map limits, so these should be carefully selected with grid cell resolution based on available memory. Second, measurement rays spanning 3D space typically involve more intersections with grid cells. The computational orders of ray casting and the inverse sensor model are proportional to the number of grid cells along the ray $n_r$, so careful consideration must be placed on sensor limits and rays per scan that can be considered. Typically, computers small enough to be carried onboard flying robots lack the processing capabilities to update massive maps quickly, particularly because the ray-by-ray updating method requires many computations in series with large measurement scans. Therefore, an onboard computer can simply stream the sensor data, and a more powerful computer can run the mapping.

Additionally, mobile robots, particularly aerial vehicles, are frequently subject to fast movements, and the time stamps for pose estimates and depth sensor readings are not guaranteed to align. For example, the most recent pose estimate and measurement scan may have significantly different time stamps while the robot is rapidly rotating. Using these together violates the assumption that the robot pose is well-known for every measurement scan, i.e., $X_t$ and $Z_t$ are given. Therefore, pose estimates and sensor readings must be paired according to their time stamps. A useful tool that easily satisfies this goal is the Message Filters package with the Robot Operating System (ROS), where stamped poses and measurement scans are synchronized using recently-received data.
